  \documentclass[a4paper,11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}

\title{CSC 2/451 - Project 1}
\author{
Yufei Du \\
\texttt{<ydu14@ur.rochester.edu>}
\and
Princeton Ferro \\
\texttt{<pferro@u.rochester.edu>}
}

\begin{document}

\maketitle

\section{Files}
All the files we modified, except for \texttt{sim\_main.cpp}, are in \texttt{verilog/}.

\section{iCache}
Implemented in \texttt{verilog/ICACHE.v} as a 32 KB read-only direct-mapped cache with 32 B block size. The current PC is used to index into the cache. $ 32KB / 32 B$ is 1024, so the index is 10 bits. Each instruction is 4-byte aligned, so the lower 2 bits can be ignored, while the 3 bits after that will be the number of instructions into the cache entry. Therefore, the lower 5 bits are the offset. This leaves us with $32-10-5 =$ 17 bits for the tag.

The 10-cycle miss penalty is achieved by an internal counter that, when nonzero and less than 10, stalls the IF and ID stages. IF and ID don't proceed if the \texttt{valid} from the iCache is negative.

Additionally, \texttt{sim\_main.cpp} was modified at lines 823 and 827 to fetch on aligned addresses.

\section{dCache}
Implemented in \texttt{verilog/DCACHE.v}, but not integrated with the rest of the MIPS pipeline because we could not get it to work. The dCache was intended to be placed after the MEM stage.

\section{Programs}
\begin{tabular}{| c | c | c | c | c |}
  \hline
  program & w/o cache & & with i-cache &\\
  \hline
  & cycles & IPC & cycles & IPC \\
  \hline
  noio & 27700 & 0.0751264 & 4272 & 0.487125\\
  file & 1307030 & 0.0728484 & 107322 & 0.88719\\
  hello & 1308960 & 0.0731153 & 106360 & 0.899821\\
  class & 1328990 & 0.0731209 & 108262 & 0.89761 \\
  sort & 1436460 & 0.0730435 & 117767 & 0.890946 \\
  fact12 & 1515830 & 0.0731085 & 123133 & 0.900002\\
  matrix & 1881850 & 0.0729527 & 153471 & 0.89454 \\
  hanoi & 2762500 & 0.0730016 & 228402 & 0.882948 \\
  ical & 2519410 & 0.0859245 & 227704 & 0.950704 \\
  fib18 & 4207410 & 0.0726692 & 316934 & 0.964709 \\
  \hline
\end{tabular}

The baseline (two left columns) is the "uncached" data, when running with the original MIPS pipeline given to us without any modifications.

\end{document}
